Shaking out the idiosyncracies

### 2020-02-11 analysis/reports/vetting.Rmd
increase macdb samples

<!-- ** I AM REALLY NOT CONFIDENT IN THE DATA HANDLING HERE, NEED TO REVISIT WHEN SHARPER ** -->

<!-- ```{r macd ctrlcomp} -->

<!-- macd_comparisons <- read.csv(here::here("working-data", "macdb_data", "orderedcomparisons.csv"), header = F, stringsAsFactors = F) -->

<!-- colnames(macd_comparisons) <- c("studyID", "control", "site") -->
<!-- macd_comparisons <- macd_comparisons %>% -->
<!--   mutate(site = as.character(site), control = as.character(control)) -->

<!-- cc_di <- all_di_macd %>% -->
<!--   filter(singletons == FALSE, treatment == "comparison") %>% -->
<!--   left_join(macd_comparisons, by = c("studyID", "site")) %>% -->
<!--   select(dat, site, skew_percentile, simpson_percentile, studyID, control) %>% -->
<!--   rename(comparison = site) %>% -->
<!--  left_join(select(all_di_macd, skew_percentile, simpson_percentile, site, singletons, nsamples), by = c("control" = "site")) %>% -->
<!--   distinct() %>% -->
<!--   rename(comp_skew = skew_percentile.x, comp_simp = simpson_percentile.x, -->
<!--          control_skew = skew_percentile.y, control_simp = simpson_percentile.y) -->

<!-- ggplot(data = cc_di, aes(x = comp_skew, y = control_skew)) + -->
<!--   geom_point(alpha = .5) + -->
<!-- #  xlim(0, 100) + -->
<!--  # ylim(0, 100) + -->
<!--   theme_bw() + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") -->

<!-- ggplot(data = cc_di, aes(x = comp_simp, y = control_simp)) + -->
<!--   geom_point(alpha = .5) + -->
<!--  xlim(0, 100) + -->
<!--  ylim(0, 100) + -->
<!--   theme_bw() + -->
<!--   geom_abline(intercept = 0, slope = 1, color = "green") -->

<!-- ``` -->

>>>>

```{r macd ctrlcomp}

macd_comparisons <- read.csv(here::here("working-data", "macdb_data", "orderedcomparisons.csv"), header = F, stringsAsFactors = F)

colnames(macd_comparisons) <- c("studyID", "control", "site")
macd_comparisons <- macd_comparisons %>%
  mutate(site = as.character(site), control = as.character(control))

cc_di <- all_di_macd %>%
  filter(singletons == FALSE, treatment == "comparison") %>%
  left_join(macd_comparisons, by = c("studyID", "site")) %>%
  select(dat, site, skew_percentile, simpson_percentile, studyID, control) %>%
  rename(comparison = site) %>%
  mutate(ctrl_skew_percentile = NA,
         ctrl_simpson_percentile = NA)

for(i in 1:nrow(cc_di)) {

  control_row <- filter(all_di_macd, site == cc_di$control[i], singletons == F) %>%
    unique()
  if(nrow(control_row) >0 ) {
    cc_di$ctrl_simpson_percentile[i] <- control_row$simpson_percentile
    cc_di$ctrl_skew_percentile[i] <- control_row$skew_percentile
  }
}

cc_di <- cc_di %>%
  mutate(simpson_change = simpson_percentile - ctrl_simpson_percentile,
         skew_change = skew_percentile - ctrl_skew_percentile)

ggplot(data = cc_di, aes(x = skew_percentile, y = ctrl_skew_percentile)) +
  geom_point(alpha = .5) +
  #  xlim(0, 100) +
  # ylim(0, 100) +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "green")

ggplot(data = cc_di, aes(x = simpson_percentile, y = ctrl_skew_percentile)) +
  geom_point(alpha = .5) +
  xlim(0, 100) +
  ylim(0, 100) +
  theme_bw() +
  geom_abline(intercept = 0, slope = 1, color = "green")

ggplot(data = cc_di, aes(x = simpson_change)) +
  geom_histogram() +
  theme_bw()

ggplot(data = cc_di, aes(x = skew_change)) +
  geom_histogram() +
  theme_bw()

ggplot(data = cc_di, aes(x = ctrl_simpson_percentile, y = simpson_change)) +
  geom_point(alpha = .3) +
  geom_hline(yintercept = 0) +
  theme_bw()

ggplot(data = cc_di, aes(x = ctrl_skew_percentile, y = skew_change)) +
  geom_point(alpha = .3) +
  geom_hline(yintercept = 0) +
  theme_bw()

print(t.test(cc_di$skew_percentile, cc_di$ctrl_skew_percentile, paired = T)
)
print(t.test(cc_di$simpson_percentile, cc_di$ctrl_simpson_percentile, paired = T)
)

print(wilcox.test(cc_di$skew_percentile, cc_di$ctrl_skew_percentile, paired = T))
print(wilcox.test(cc_di$simpson_percentile, cc_di$ctrl_simpson_percentile, paired = T))

Change is going to be bounded at 100 and 0: you can't go up or down from there. (Another argument for increasing the number of samples?)
